\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-04-15 paper.qmd
code tidied up some more and caption tidied

umap_scurve_scaled <- gen_scaled_data(data = umap_scurve, x = "UMAP1", 
                                      y = "UMAP2", hr = 2/sqrt(3))

# ar <- umap_scurve |>
#   summarise_at(c("UMAP1", "UMAP2"), 
#     function(x) max(x)-min(x)) |>
#   mutate(ar = UMAP2/UMAP1) |>
#   pull(ar) # This is a bit cumbersome too
# hr <- 2/sqrt(3)
# umap_scurve_scaled <- umap_scurve |>
#   mutate_at(c("UMAP1", "UMAP2"), 
#     function(x) (x - min(x))/(max(x)-min(x))) |> # can be no missings
#   mutate(UMAP2 = UMAP2 * ceiling(ar/hr) * hr)

  #as.data.frame(do.call(cbind, gen_scaled_data(data = umap_scurve, 
  #                                  x = "UMAP1", y = "UMAP2"))) |>
  #rename(c("UMAP1" = "scaled_UMAP1", 
  #                "UMAP2" = "scaled_UMAP2")) |>
  #mutate(ID = 1:NROW(umap_scurve))

nldr_scurve <- umap_scurve_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  #coord_fixed() +
  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
  interior_annotation("a", c(0.08, 0.93))
```

@@

## Hexagonal bining to have regular hexagons
hb_obj_scurve <- hex_binning(data = umap_scurve_scaled, x = "UMAP1",
                      y = "UMAP2", num_bins_x = num_bins_x_scurve,
                      num_bins_y = num_bins_y_scurve, x_start = x_start_scurve,
                      y_start = y_start_scurve, buffer_x = buffer_x_scurve,
                      buffer_y = buffer_y_scurve, hex_size = hex_size_scurve,
                      col_start = "UMAP")

## Data set with all possible centroids in the hexagonal grid
all_centroids_df <- hb_obj_scurve$centroids

@@

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_scurve$std_cts
df_bin_centroids <- extract_hexbin_centroids(centroids_df = all_centroids_df, 
                                             counts_df = counts_df)

@@

hex_grid_scurve <-  ggplot(data = hex_grid_with_counts, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
  geom_point(data = umap_scurve_scaled, aes(x = UMAP1, y = UMAP2), alpha = 0.5) +
  #coord_fixed() +
  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
  interior_annotation("b", c(0.08, 0.93))
```

@@

hex_grid_nonempty_scurve <-  ggplot(data = hex_grid_nonempty, aes(x = x, y = y)) +
  geom_polygon(color = "black", aes(group = hex_poly_id), fill = "#ffffff") +
  geom_point(data = df_bin_centroids, aes(x = c_x, y = c_y), color = "#33a02c") +
  #coord_fixed() +
  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
  interior_annotation("c", c(0.08, 0.93)) 
```

@@

```{r}
## Triangulate bin centroids
tr1_object_scurve <- tri_bin_centroids(df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df_scurve <- gen_edges(tri_object = tr1_object_scurve)

## Compute 2D distances
distance_scurve <- cal_2d_dist(tr_coord_df = tr_from_to_df_scurve, 
                             start_x = "x_from", start_y = "y_from", 
                             end_x = "x_to", end_y = "y_to", 
                             select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_scurve <- find_lg_benchmark(distance_edges = distance_scurve, distance_col = "distance")

trimesh_removed_scurve <- vis_rmlg_mesh(distance_edges = distance_scurve, 
                                        benchmark_value = benchmark_scurve, 
                                        tr_coord_df = tr_from_to_df_scurve, 
                                        distance_col = "distance") +
  xlim(c(-0.3, 1.3)) + ylim(c(-0.3, 2.6)) +
  interior_annotation("d", c(0.08, 0.93)) 
```

```{r}
#| label: fig-NLDR-scurve
#| echo: false
#| fig-cap: "Key steps for constructing the model on the NLDR layout ($k=2$) of the S-curve data: (a) obtain UMAP data, (b) hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d) triangulate bin centroids."
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%

>>>>

umap_scurve_scaled <- gen_scaled_data(
  data = umap_scurve, 
  x = "UMAP1", 
  y = "UMAP2", 
  hr = 2/sqrt(3))

sc_ltr_pos <- c(0.08, 0.96)
sc_xlims <- c(-0.25, 1.25)
sc_ylims <- c(-0.25, 2.59)
nldr_scurve <- umap_scurve_scaled |> 
  ggplot(aes(x = UMAP1, y = UMAP2)) + 
  geom_point(alpha=0.5, colour="#000000", size = 0.5) +
  #coord_fixed() +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("a", sc_ltr_pos)
```

@@

## Hexagonal bining to have regular hexagons
hb_obj_scurve <- hex_binning(
  data = umap_scurve_scaled, 
  x = "UMAP1",
  y = "UMAP2", 
  num_bins_x = num_bins_x_scurve,
  num_bins_y = num_bins_y_scurve, 
  x_start = x_start_scurve,
  y_start = y_start_scurve, 
  buffer_x = buffer_x_scurve,
  buffer_y = buffer_y_scurve, 
  hex_size = hex_size_scurve,
  col_start = "UMAP")

## Data set with all centroids
all_centroids_df <- hb_obj_scurve$centroids

@@

## To obtain the standardise counts within hexbins
counts_df <- hb_obj_scurve$std_cts
df_bin_centroids <- extract_hexbin_centroids(
  centroids_df = all_centroids_df, 
  counts_df = counts_df)

@@

hex_grid_scurve <- ggplot(
  data = hex_grid_with_counts, 
  aes(x = x, y = y)) +
  geom_polygon(color = "black", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = umap_scurve_scaled, 
             aes(x = UMAP1, y = UMAP2), 
             alpha = 0.5) +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("b", sc_ltr_pos)
```

@@

hex_grid_nonempty_scurve <-  ggplot(
  data = hex_grid_nonempty, 
  aes(x = x, y = y)) +
  geom_polygon(color = "black", 
               aes(group = hex_poly_id), 
               fill = "#ffffff") +
  geom_point(data = df_bin_centroids, 
             aes(x = c_x, y = c_y), 
             color = "#33a02c") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("c", sc_ltr_pos) 
```

@@

```{r}
## Triangulate bin centroids
tr1_object_scurve <- tri_bin_centroids(
  df_bin_centroids, x = "c_x", y = "c_y")
tr_from_to_df_scurve <- gen_edges(
  tri_object = tr1_object_scurve)

## Compute 2D distances
distance_scurve <- cal_2d_dist(
  tr_coord_df = tr_from_to_df_scurve, 
  start_x = "x_from", 
  start_y = "y_from", 
  end_x = "x_to", 
  end_y = "y_to", 
  select_vars = c("from", "to", "distance"))

## To find the benchmark value
benchmark_scurve <- find_lg_benchmark(
  distance_edges = distance_scurve, 
  distance_col = "distance")

trimesh_removed_scurve <- vis_rmlg_mesh(
  distance_edges = distance_scurve, 
  benchmark_value = benchmark_scurve, 
  tr_coord_df = tr_from_to_df_scurve, 
  distance_col = "distance") +
  xlim(sc_xlims) + ylim(sc_ylims) +
  interior_annotation("d", sc_ltr_pos) 
```

```{r}
#| label: fig-NLDR-scurve
#| echo: false
#| fig-cap: "Key steps for constructing the model on the UMAP layout ($k=2$) of the S-curve data: (a) data, (b) hexagon bins, (c) bin centroids, and (d) triangulated centroids."
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%

### 2024-04-15 paper.tex
Fix issues
code tidied up some more and caption tidied

\caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
the NLDR layout (\(k=2\)) of the S-curve data: (a) obtain UMAP data, (b)
hexagonating UMAP data, (c) obtain hexagonal bin centroids, and (d)
triangulate bin centroids.}

@@

\begin{itemize}
\tightlist
\item
  NLDR view used to illustrate clusters
\item
  Use our method to assess is it a reasonable representation
\item
  Demonstrate that it is not
\item
  Illustrate how to use out method to get a better representation
\end{itemize}

>>>>

\caption{\label{fig-NLDR-scurve}Key steps for constructing the model on
the UMAP layout (\(k=2\)) of the S-curve data: (a) data, (b) hexagon
bins, (c) bin centroids, and (d) triangulated centroids.}

@@

\begin{itemize}
\tightlist
\item
  NLDR view used to illustrate clusters
\item
  Use our method to assess is it a reasonable representation
\item
  Demonstrate that it is not
\item
  Illustrate how to use our method to get a better representation
\end{itemize}
